---
title: "Logistic Regression Covariance Exercise"
author: "C. Durso"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
library(tidyverse)
library(MASS)
```

## Purpose

This illustrates the use of the covariance matrix of a logistic regression to test generate multiple $\hat{\boldsymbol{\beta}}$ values. These can be used to estimate the variability of the probability of the outcome given explanatory variables.

## Data

```{r}
dat<-read.csv("SAheart.data")
dat$row.names<-NULL
dat$famhist<-(dat$famhist=="Present")*1
```

## Model

```{r}
fmla.backward.best<-as.formula("chd ~ 1 + age + ldl + famhist + tobacco + I(typea^2) + I(age^2) + 
    I(sbp^2)")
```

## Sample from $\hat{\boldsymbol{\beta}}$ distribution

```{r}

```

## Model "chd" probabilities

Using the data from the 15th observation and the collection of sets of reasonable coefficients, model possible probabilities of "chd" for case 15.

```{r}
# original estimate


# vector of observations
# You may find model.matrix useful with this model.


# matrix with columns equal to collections of coefficients


# values of the linear predictor


# values of the logistic function of the linear predictor,
# aka probabilities

# .025,.5,.975 quantiles of simulated probabilities

```

