---
title: "Multiple Regression Covariance Exercise"
author: "C. Durso"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
library(tidyverse)
```

## Purpose

This illustrates the use of the covariance matrix of a multiple regression to test the hypothesis of equality of two coefficients.

## Data

The data are body measurements, age, and gender for a collection of physically active individuals.

Measurements were initially taken by Grete 
Heinz and Louis J. Peterson at San Jose State University and at the 
U.S. Naval Postgraduate School in Monterey, California. Later, 
measurements were taken at dozens of California health and fitness 
clubs by technicians under the supervision of one of these authors.

Downloaded 9/5/2019 from http://jse.amstat.org/datasets/body.dat.txt

The variable names are supplied from the documentation at 
http://jse.amstat.org/datasets/body.txt

```{r}
dat<-read.table("body.dat.txt")
nam<-c("biacromial.diam","biiliac.diam", "bitrochanteric.diam","chest.depth","chest.diam", "elbow.diam","wrist.diam", "knee.diam","ankle.diam","shoulder.girth", "chest.girth", 
 "waist.girth", "Navel.girth", "hip.girth","Thigh.girth", 
"Bicep.girth", "Forearm.girth", "Knee.girth", "Calf.girth",
"Ankle.girth", "Wrist.girth", "Age", "Weight", "height","Gender")
nam<-str_to_lower(nam)
names(dat)<-nam
dat<-dat[dat$gender==0,] # restrict to female cases
```

## Forward model

The body measurement data was used to estimate weight based on the remaining measurements. These measurements address the volume of the individual. The measurements selected by a forward model on training data using the AIC are I(hip.girth^2),I(chest.girth^2), I(calf.girth^2), height, I(waist.girth^2), I(knee.girth^2),  I(thigh.girth^2), chest.girth, I(forearm.girth^2), wrist.diam,  biiliac.diam. (Somewhat surprisingly, all interactions of height with the remaining measurements are dropped. While the interactions may not show during the trace, they are considered.)

Fit this model on the full data set with scaled measurements.
       

```{r}
dat<-dplyr::select(dat,-gender)
dat.scale<-data.frame(scale(dat))
dat.scale$weight<-dat$weight
```
