---
title: "Poisson Best Subsets Regression"
author: "C. Durso"
output:
  pdf_document: default
  html_document:
    df_print: paged
---
```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
library(tidyverse)
library(bestglm)
```

## Introduction

The data set "cells.csv" from Crawley, Statistics: an Introduction using R gives the number of abnormal cells in samples from individuals with additional data provided. The goal here is to fit a best subset model according to BIC for Poisson regression model on the training data set with all pairwise interactions and to assess the performance using log likelihood on a test set.


## Data

```{r, cache=TRUE}

dat<-read.csv("cells.csv")
X<-model.matrix(as.formula("cells~smoker+sex+age+weight"),
                data=dat)
X<-X[,-1]
Y<-dat$cells

X<-cbind(X,Y)
X<-data.frame(X)
names(X)[ncol(X)]<-"cells"

set.seed(3456787)
train.ind<-sample(1:nrow(dat),round(.7*nrow(dat)))
test.ind<-setdiff(1:nrow(dat),train.ind)

dat.train<-X[train.ind,]
dat.test<-X[test.ind,]
```


### Best subset model, BIC

Fit a best subset model according to the BIC on the training data. Due to algorithmic limitations, the models use just the original variables.

```{r}
set.seed(3456787)
poissbest<-bestglm(dat.train, IC="BIC", family=poisson)
m.poiss<-poissbest$BestModel
summary(m.poiss)
```

### Log Likelihood

Calculate the log likelihood of this model on the training data.

```{r}

```











