---
title: "Poisson Regression Forward Model"
author: "C. Durso"
output:
  pdf_document: default
  html_document:
    df_print: paged
---
```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
library(tidyverse)
library(MASS)
```

## Introduction

The data set "cells.csv" from Crawley, Statistics: an Introduction using R gives the number of abnormal cells in samples from individuals with additional data provided. The goal here is to fit a forward Poisson regression model to the data set with all pairwise interactions and to assess the performance on a test set using log likelihood.


## Data

```{r, cache=TRUE}

dat<-read.csv("cells.csv")
set.seed(3456787)
train.ind<-sample(1:nrow(dat),round(.7*nrow(dat)))
test.ind<-setdiff(1:nrow(dat),train.ind)

dat.train<-dat[train.ind,]
dat.test<-dat[test.ind,]
```

## Forward Model

Fit a forward model using AIC as the stopping criterion

```{r}
m<-glm(cells~1,data=dat.train,family="poisson")
m.forward<-step(m,scope=~(smoker+sex+age+weight)^2,trace=0)
summary(m.forward)
```

### Log Likelihood

Calculate the log likelihood on the test data.

```{r}
pred<-predict(m.forward,dat.test,type="response")

# your computation of log likelihood here


```

