---
title: "Poisson Regression Example"
author: "C. Durso"
output:
  pdf_document: default
  html_document: default
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
library(tidyverse)
```

## Introduction

Regression for count data can use a non-Gaussian error process. The family of Poisson distributions is a one-parameter family of distributions with outcomes in the positive integers. The Poisson distribution with parameter $\lambda$ has the probability mass function $p(k)=\frac{\lambda^ke^{-\lambda}}{k!}$. 

## Data

The data "clusters.csv" are from *Statistics, An introduction using R*. They represent the number of reported cancer cases per year per clinic, together with the distance from a nuclear plant to the clinic in km.

```{r}
dat<-read.csv("../../text_data/clusters.csv")
qplot(data=dat,Distance,Cancers)
```

## Poisson model
```{r}
m<-glm(Cancers~Distance,data=dat,family="poisson")
summary(m)
```




## Model Selection

Many of the model selection methods discussed for linear regression are available for Poisson regression (though fewer for quasipoisson). 
